---
title: "A Bayesian workflow with mdir"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Bayesian workflow with mdir}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 3.5)
options(mdir.quiet = TRUE)
```

This vignette walks through the steps recommended before trusting an `mdir`
analysis: choosing and checking priors, fitting several chains, checking
convergence and checking that the fitted model reproduces the data. The data are
simulated, and the chains are far shorter than you would use in practice.

```{r data}
library(mdir)
set.seed(1)

N <- 60
truth <- rep(1:2, each = N / 2)
make_view <- function(P, separation) {
  x <- matrix(rnorm(N * P, mean = ifelse(truth == 1, 0, separation)), N, P)
  rownames(x) <- seq_len(N)
  x
}
X <- list(make_view(3, 3), make_view(2, 2.5))

# Data may contain missing values; they are imputed within the sampler
X[[1]][sample(length(X[[1]]), 15)] <- NA
mean(is.na(X[[1]]))
```

## Priors

The MDI-level priors (concentration of the component weights and the strength of
association between views) are set with `mdiPrior()`, the priors of the component
densities with `densityPrior()`. See their help pages for what each prior means
and how to choose it.

```{r priors}
mdiPrior()
densityPrior()
```

A prior predictive check shows the data and the cluster agreement between views
that the priors imply before the data are seen.

```{r prior-predictive}
types <- c("MVN", "G")
K <- c(4, 4)
prior_sims <- simulatePriorPredictive(X, types, K = K, n_datasets = 30)
plotPredictiveCheck(X, prior_sims, view = 1, column = 1)
```

## Fitting and convergence

```{r fit}
chains <- runMCMCChains(X, n_chains = 3, R = 1500, thin = 5, types = types, K = K)
assessConvergence(chains, burn = 750)
```

The diagnostics are the rank-normalised split R-hat and bulk and tail effective
sample sizes of Vehtari et al. (2021), computed for quantities that do not depend
on how the clusters are labelled. Short chains such as these usually fail the
criteria; use more iterations for real analyses.

## Posterior predictive check

```{r posterior-predictive}
post_sims <- simulatePosteriorPredictive(chains, X, n_draws = 50, burn = 750)
plotPredictiveCheck(X, post_sims, view = 1, column = 1)
predictiveCheck(X, post_sims, function(m) sd(m[, 1], na.rm = TRUE), view = 1)$p_value
```

Posterior predictive p-values are conservative, because the data are used both to
fit and to check the model: values near 0 or 1 indicate misfit, while values
in the middle do not prove that the model fits.

## Point estimates

```{r predict}
pooled <- predictFromMultipleChains(chains, burn = 750)
table(pooled$pred[[1]], truth)
```

## References

Vehtari, A., Gelman, A., Simpson, D., Carpenter, B. and Burkner, P.-C. (2021).
Rank-normalization, folding, and localization: an improved R-hat for assessing
convergence of MCMC. *Bayesian Analysis*, 16(2), 667-718.
