---
title: "A Bayesian workflow with mdir"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Bayesian workflow with mdir}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 3.5)
options(mdir.quiet = TRUE)
```

This vignette walks through the steps recommended before trusting an `mdir`
analysis: choosing and checking priors, fitting several chains, checking
convergence and checking that the fitted model reproduces the data. The data are
simulated, and the chains are far shorter than you would use in practice.

```{r data}
library(mdir)
set.seed(1)

N <- 60
truth <- rep(1:2, each = N / 2)
make_view <- function(P, separation) {
  x <- matrix(rnorm(N * P, mean = ifelse(truth == 1, 0, separation)), N, P)
  rownames(x) <- seq_len(N)
  x
}
X <- list(make_view(3, 3), make_view(2, 2.5))

# Data may contain missing values; they are imputed within the sampler
X[[1]][sample(length(X[[1]]), 15)] <- NA
mean(is.na(X[[1]]))
```

## Priors

The MDI-level priors (concentration of the component weights and the strength of
association between views) are set with `mdiPrior()`, the priors of the component
densities with `densityPrior()`. See their help pages for what each prior means
and how to choose it.

```{r priors}
mdiPrior()
densityPrior()
```

A prior predictive check shows the data and the cluster agreement between views
that the priors imply before the data are seen.

```{r prior-predictive}
types <- c("MVN", "G")
K <- c(4, 4)
prior_sims <- simulatePriorPredictive(X, types, K = K, n_datasets = 30)
plotPredictiveCheck(X, prior_sims, view = 1, column = 1)
```

## Fitting and convergence

`fitMDI()` runs the chains and checks convergence before returning. It
reports progress as each chain finishes (silenced here with
`options(mdir.quiet = TRUE)`; use `verbose = TRUE` to see it).

```{r fit}
chains <- fitMDI(X, n_chains = 3, R = 1500, thin = 5, types = types, K = K,
                 burn = 750, verbose = FALSE)
chains
```

Printing the result gives a short report rather than the sampled arrays.
`summary()` adds a table per chain and the full convergence table:

```{r summary}
summary(chains)
```

The diagnostics are the rank-normalised split R-hat and bulk and tail effective
sample sizes of Vehtari et al. (2021), computed for quantities that do not depend
on how the clusters are labelled. Short chains such as these usually fail the
criteria; use more iterations for real analyses.

`plot()` draws the same quantities. Trace plots come first: chains that overlap
and show no trend after the burn in (dashed line) are consistent with
convergence. `type = "density"` overlays the posterior of each quantity by chain,
and `type = "rhat"` plots the diagnostics against their thresholds.

```{r plot-trace, fig.width = 7, fig.height = 4}
plot(chains, burn = 750)
```

```{r plot-rhat, fig.width = 7, fig.height = 3.5}
plot(chains, type = "rhat", burn = 750)
```

## Posterior summaries and the draws format

`as_draws()` converts the chains to the format of the \pkg{posterior} package,
so the Stan workflow applies: `summarise_draws()` for means, quantiles, Rhat and
effective sample sizes, `subset_draws()`, and \pkg{bayesplot} functions such as
`mcmc_trace()` and `mcmc_pairs()`. A mixture has no parameters that mean the same
thing in every draw, so the variables are the quantities that do not depend on the
labelling: log-likelihoods, the concentrations, `phi`, the number of occupied
components and the pooled hyperparameters. `weights = TRUE` adds the sorted
component weights and `allocations = TRUE` the item allocations.

```{r draws, eval = requireNamespace("posterior", quietly = TRUE)}
draws <- posterior::as_draws(chains, burn = 750)
posterior::summarise_draws(draws)
```

For leave-one-out cross-validation or WAIC, `pointwiseLogLik()` gives the
per-item log-likelihoods that \pkg{loo} takes.

## Posterior predictive check

```{r posterior-predictive}
post_sims <- simulatePosteriorPredictive(chains, X, n_draws = 50, burn = 750)
plotPredictiveCheck(X, post_sims, view = 1, column = 1)
predictiveCheck(X, post_sims, function(m) sd(m[, 1], na.rm = TRUE), view = 1)$p_value
```

Posterior predictive p-values are conservative, because the data are used both to
fit and to check the model: values near 0 or 1 indicate misfit, while values
in the middle do not prove that the model fits.

## Point estimates

```{r predict}
pooled <- predictFromMultipleChains(chains, burn = 750)
table(pooled$pred[[1]], truth)
```

The posterior similarity matrix shows how sure the partition of a view is, and
the fusion probabilities show, for each item, how often it takes the same
component in two views. Fusion probabilities compare component indices, so they
describe fusion of clusters only if the indices are aligned between views.

```{r plot-psm, fig.width = 5, fig.height = 4}
plot(chains, type = "psm", view = 1, burn = 750)
```

```{r plot-fusion, fig.width = 5, fig.height = 3.5}
plot(chains, type = "fusion", burn = 750)
```

## New items

`predictMDI()` evaluates items that were not in the fitted data at every saved
draw, using the draw's weights, association parameters and component parameters.
It returns the log posterior predictive density of each new item and the
probability that it shares a component with each fitted item. (For a semi-supervised
view it also returns class probabilities; labels of other views are only defined up
to permutation, so averaging them across draws would be meaningless.) A new item may
be missing entries or a whole view.

```{r new-items}
new_items <- list(
  matrix(c(0.1, -0.2, 0.3, 2.9, 3.1, 3.2), 2, 3, byrow = TRUE),
  matrix(c(0.0, 0.2, 2.5, NA), 2, 2, byrow = TRUE)
)
new_pred <- predictMDI(chains, X, new_items, burn = 750, coclustering = TRUE)
round(new_pred$log_predictive, 1)

# Probability that each new item is in the same cluster as the first and last
# fitted items, in view 1
round(new_pred$coclustering[[1]][, c(1, N)], 2)
```

## Model comparison

Fit with `save_pointwise = TRUE` to record the log-likelihood of each item at each
saved draw. It sums over the cluster assignments of the item in every view, so it is
the likelihood of the whole model, coupling between views included (the
`observed_likelihood` trace treats each view separately and ignores the coupling).
`pointwiseLogLik()` returns it as the draws by items matrix that `loo::waic()` and
`loo::loo()` take.

```{r pointwise, eval = FALSE}
fit <- fitMDI(X, n_chains = 4, R = 1500, thin = 5, types = types, K = K,
              save_pointwise = TRUE)
ll <- pointwiseLogLik(fit, burn = 750)
loo::waic(ll)
```

## Running chains in parallel

Chains are independent, so `n_cores` runs them on separate cores. The default is one
core. Each chain has its own random number stream, so results are reproducible
from `set.seed()` and do not depend on the number of cores. While a package is
being checked, at most two cores are used, as CRAN requires.

```{r parallel, eval = FALSE}
chains <- fitMDI(X, n_chains = 4, R = 5000, thin = 5, types = types, K = K,
                 n_cores = 4)
```

## References

Vehtari, A., Gelman, A., Gabry, J. (2017). Practical Bayesian model evaluation
using leave-one-out cross-validation and WAIC. *Statistics and Computing*, 27,
1413-1432.

Vehtari, A., Gelman, A., Simpson, D., Carpenter, B. and Burkner, P.-C. (2021).
Rank-normalization, folding, and localization: an improved R-hat for assessing
convergence of MCMC. *Bayesian Analysis*, 16(2), 667-718.
