---
title: "Multimodal clustering with MDI: an end-to-end workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multimodal clustering with MDI: an end-to-end workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
options(mdir.quiet = TRUE)
have_posterior <- requireNamespace("posterior", quietly = TRUE)
have_loo <- requireNamespace("loo", quietly = TRUE)
```

This vignette follows one analysis from question to report, using the steps of a
Bayesian workflow (Gelman et al., 2020): say what the model assumes, check the
priors against what you know, fit with several chains, diagnose, check the fit,
test sensitivity, and only then interpret. It is about what is specific to
multiple dataset integration (MDI; Kirk et al., 2012), and to this implementation.
The data are simulated so that the answer is known, which lets the last sections
check whether the method finds what is there. The chains are shorter than you
would use for a real analysis.

```{r libs}
library(mdir)
set.seed(2026)
```

## What MDI does, and what it does not

MDI clusters several datasets (modalities) of the *same items*. It does **not**
estimate one global clustering. Each modality has its own mixture model, and
therefore its own partition of the items. The partitions are coupled: the prior
probability that item $n$ takes components $(c_{n1}, \dots, c_{nL})$ in the $L$
views is proportional to

$$\prod_{l} w_{l c_{nl}} \prod_{l < m} \bigl(1 + \phi_{lm}\,\mathbf 1[c_{nl} = c_{nm}]\bigr).$$

Here $w_{lk}$ are the component weights of view $l$, and $\phi_{lm} \ge 0$ measures
how much views $l$ and $m$ share structure: an item that takes the same component
index in both views is $1 + \phi_{lm}$ times more probable than one that does not,
before the data are seen. With $\phi_{lm} = 0$ the views are independent mixtures.
Three consequences shape every step below.

1. **The questions MDI answers are about views and items, not about one set of
   clusters.** What is the partition in each view? How strongly are views
   associated ($\phi$)? Which individual items are placed consistently in two
   views (their *fusion probability*), and which are not? Borrowing of strength
   works through $\phi$: an item that is ambiguous in one view is pulled towards
   the component that its neighbours in the other view take.
2. **Association is partial and pairwise.** Views 1 and 2 can share structure
   while view 3 is unrelated, and a single item can follow the shared structure
   in one view and not in another. A model that forces one partition would
   mis-describe this; MDI lets the data say how much is shared.
3. **Component indices carry meaning across views.** The coupling compares the
   index $c_{nl}$ with $c_{nm}$, so "cluster 2 of view 1" and "cluster 2 of view
   2" are the same cluster only to the extent that the sampler has aligned them.
   The sampler has a move that exchanges two components within a view to improve
   alignment, and the fusion probabilities depend on it. Anything that compares
   indices across views (fusion probabilities, `agreement[l,m]`) inherits this;
   quantities such as `phi`, the number of occupied components and the
   adjusted Rand index between views' partitions (below) do not.

Two further properties of the model matter for how you read the output. The number
of components $K$ is an *upper bound*: with a sparse prior on the weights the
posterior empties surplus components (Rousseau and Mengersen, 2011), so choose $K$
generously and look at the occupied components, not at $K$. And because a mixture
posterior is invariant to relabelling, parameters of individual components have no
fixed meaning from one draw to the next. Everything that is monitored here is
therefore either label-invariant or explicitly aligned.

## Data and the question

We simulate $N = 120$ items measured in three modalities.

* View 1 (continuous, 2 features) has three clusters.
* View 2 (continuous, 3 features) has the same three clusters for most items. For
  18 items, chosen at random, the cluster is re-drawn at random: these items do
  **not** follow the shared structure.
* View 3 (four binary features) has two clusters that are unrelated to the other
  views.

The questions are: do views 1 and 2 share structure and view 3 not? Which items in
views 1 and 2 break the pattern? What are the partitions?

```{r data}
N <- 120
z1 <- rep(1:3, each = N / 3)
z2 <- z1
broken <- sort(sample(N, 18))
z2[broken] <- sample(1:3, 18, replace = TRUE)
z3 <- sample(1:2, N, replace = TRUE)

mu1 <- rbind(c(-3, 0), c(0, 3), c(3, 0))
mu2 <- rbind(c(-2, -2, 0), c(2, -2, 2), c(0, 3, -2))
p3 <- rbind(c(0.8, 0.2, 0.2, 0.7), c(0.2, 0.8, 0.7, 0.2))

X <- list(
  mu1[z1, ] + matrix(rnorm(2 * N), N, 2),
  mu2[z2, ] + matrix(rnorm(3 * N), N, 3),
  matrix(rbinom(4 * N, 1, p3[z3, ]), N, 4)
)
for (v in 1:3) rownames(X[[v]]) <- sprintf("item%03d", 1:N)
types <- c("MVN", "G", "C")   # full covariance, diagonal covariance, categorical
```

Items must be in the same order in every modality (this is checked through the
row names). Missing entries (`NA`) are allowed and are imputed inside the sampler;
an item missing from a whole modality contributes only through the others.
Categorical data are coded `0, 1, 2, ...`.

A fast look at what the views have in common, before any modelling, is the cross
table of two views' simple clusterings. The sampler will do this properly, but it
guards against a data-handling mistake (rows in different orders, for instance)
that no later diagnostic would catch.

```{r look}
km <- lapply(1:2, function(v) stats::kmeans(X[[v]], 3, nstart = 10)$cluster)
table(view1 = km[[1]], view2 = km[[2]])
```

## Priors, and what they imply

The priors specific to MDI are on the concentration of each view's weights
(`mass`), and on the association parameters `phi`. Printing the defaults shows
them:

```{r prior}
mdiPrior()
```

The component densities have conjugate priors whose location and scale are set
from the data (see `?mdiPrior`), with the variance scale pooled across
components (`?densityPrior`). That uses the data twice, weakly; it is a deliberate
trade-off, standard for mixtures, and worth stating in a report.

**The sparsity check.** Fitting with `K = 6` components per view and the default
prior on `mass` prints a message for the views where the prior on the weights is
on the side of the Rousseau-Mengersen threshold at which superfluous components
*duplicate* occupied ones instead of emptying. The result is asymptotic and does
not cover a random concentration, so read it as guidance on where the prior sits.

```{r sparsity-default}
K <- c(6, 6, 6)
old <- options(mdir.quiet = FALSE)          # the messages are silenced in this vignette
invisible(callMDI(X, R = 10, thin = 5, types = types, K = K))
options(old)
```

Lowering the prior on `mass` makes the weights sparse and removes the message; the
default on `phi` is kept.

```{r sparsity}
prior <- mdiPrior(mass_shape = 2, mass_rate = 0.5)
old <- options(mdir.quiet = FALSE)
invisible(callMDI(X, R = 10, thin = 5, types = types, K = K, prior = prior))
options(old)
prior
```

**Prior predictive check.** The prior on `phi` is hard to judge directly. What
matters is what it implies about the *data*, so simulate from the prior and look.
The implied agreement between two views' partitions is a label-invariant summary
we can compute for every prior draw with the adjusted Rand index (ARI), defined
here once and used throughout.

```{r ari}
ari <- function(a, b) {
  t <- table(a, b)
  n <- sum(t)
  s <- function(x) sum(choose(x, 2))
  expected <- s(rowSums(t)) * s(colSums(t)) / choose(n, 2)
  maximum <- 0.5 * (s(rowSums(t)) + s(colSums(t)))
  if (maximum == expected) return(NA_real_)
  (s(t) - expected) / (maximum - expected)
}
```

```{r prior-predictive}
prior_sims <- simulatePriorPredictive(X, types, K = K, n_datasets = 200, prior = prior)
prior_ari <- vapply(prior_sims$parameters, function(p) ari(p$labels[, 1], p$labels[, 2]),
                    numeric(1))
summary(prior_ari)
hist(prior_ari, breaks = 20, main = "Prior: ARI between the partitions of views 1 and 2",
     xlab = "Adjusted Rand index", col = "grey85", border = "white")
```

The prior spreads over weak and strong association rather than asserting either,
which is what we want to say before seeing the data. If you expect little sharing
(for instance in a screen across unrelated assays), lower the mean of the `phi`
prior with `phi_shape` and `phi_rate`, and re-run this check. The same simulations
give datasets to compare with yours:

```{r prior-data-check}
plotPredictiveCheck(X, prior_sims, view = 1, column = 1)
```

## Fitting

Run several chains from dispersed starting points. Each chain starts from labels
drawn from a stick-breaking prior, so the starts differ without any action from you.
`fitMDI()` runs the chains and checks convergence before it returns. We start with
plain chains, which are cheap, and four of them.

```{r fit-plain}
set.seed(105)
fit_plain <- fitMDI(X, n_chains = 4, R = 6000, thin = 10, types = types, K = K,
                    prior = prior, burn = 3000, verbose = FALSE)
```

## Convergence for a mixture model

Mixture posteriors are multimodal in the labelling and often in the partition, so
the usual rule of "look at Rhat" needs three adjustments.

* **Monitor label-invariant quantities.** `fitMDI()` monitors log-likelihoods, the
  concentrations, `phi`, the number of occupied components, the pooled
  hyperparameters, and the agreement between views (which relies on the alignment
  of components across views). Rhat of a component's mean would be meaningless
  because components swap.
* **Expect the concentration to mix slowest.** `mass` is informed only by how many
  components are occupied, and it is a nuisance parameter. A low effective sample
  size there does not hurt the questions asked, but say so. What matters is that
  `phi` and the partitions are stable.
* **Chains that agree prove little; chains that disagree prove a lot.** Chains that
  start in different modes of the partition and stay there disagree, which Rhat
  catches. Chains that all start in the same basin and stay agree, and nothing in the
  chains can catch that.

```{r convergence}
plot(fit_plain, type = "trace", burn = 3000)
```

In the trace plots look at the log-likelihoods and the occupied components, which
are the quantities that show a chain sitting in a different mode.

```{r convergence-rhat}
plot(fit_plain, type = "rhat")
```

The summary by chain shows where a disagreement comes from. If one chain has a
higher log-likelihood *and* more occupied components than the others, it is sitting in a different
mode of the partition.

```{r per-chain}
per_chain <- function(f, burn) {
  t(vapply(unclass(f)[seq_along(f)], function(ch) {
    keep <- seq(floor(burn / ch$thin) + 2, nrow(ch$mass))
    c(joint_loglik = mean(ch$joint_likelihood[keep]),
      occupied_1 = mean(apply(ch$allocations[keep, , 1], 1, function(z) length(unique(z)))),
      occupied_2 = mean(apply(ch$allocations[keep, , 2], 1, function(z) length(unique(z)))),
      phi_12 = mean(ch$phis[keep, 1]))
  }, numeric(4)))
}
round(per_chain(fit_plain, 3000), 2)
```

This is not a quirk of the example. In MDI, the coupling makes some modes sticky:
once a cluster is split into two in *both* of two strongly associated views (or an
item forms its own small cluster in both), a move that merges the clusters in one
view alone makes the allocations disagree, which the coupling penalises heavily, so
the chain stays where it is. Such modes have higher likelihood and more
components, and they are not necessarily where the posterior mass is. When we
repeated this analysis with other random seeds, plain chains of this length
disagreed in a sizeable share of the runs. Pooling chains in which one is stuck would
give the wrong answer, and dropping the stuck chain would be a choice made after seeing
the results. The options in increasing order of cost are: more iterations;
`split_merge = 2`, which re-allocates items between two components of a view; and
parallel tempering, `betas = ptLadder(4)`, which lets a chain at a high temperature
cross barriers that the posterior chain cannot. For a posterior with several
well-separated modes, `smcMDI()` and `smcReplicates()` give weighted ensembles that do
not depend on the chains mixing, at the cost of weights that can degenerate. None
of these is guaranteed to mix; they reduce the chance that chains stay apart, and the
diagnostics are how you find out whether they did.

We rerun with tempering and split-merge moves, four chains again. This is about ten
times as costly per iteration.

```{r fit}
set.seed(205)
burn <- 1500
fit <- fitMDI(X, n_chains = 4, R = 3000, thin = 10, types = types, K = K, prior = prior,
              betas = ptLadder(4, 0.1), split_merge = 2, burn = burn,
              save_pointwise = TRUE, verbose = FALSE)
fit
round(per_chain(fit, burn), 2)
```

```{r convergence-final}
plot(fit, type = "trace", burn = burn)
```

Now the four chains sit together. The effective sample sizes of the quantities of
interest are the thing to read in the table. `as_draws()` puts the chains in the format of the `posterior` package,
so its whole toolkit (and `bayesplot`, which takes the same objects) applies:

```{r draws, eval = have_posterior}
draws <- posterior::as_draws(fit, burn = burn)
posterior::summarise_draws(draws)
```

For the quantities we interpret (the log-likelihoods, `phi` and the occupied
components) Rhat is close to one and the bulk effective sample sizes are in the
hundreds. The concentrations `mass`, which we do not interpret, mix slowest and do
not meet the criteria, as expected. Hundreds of effective draws are enough for medians
and intervals, and thin for tail probabilities. For a report you would run longer,
until the criteria are met for everything you interpret, and say which quantities you
did not require to meet them and why. The tempered run is not guaranteed to succeed:
in repeats of this analysis with other seeds it met the criteria for these quantities
in most runs and not all, so the diagnostics must be read every time.

## Posterior summaries

### How much do the views share?

`phi[l,m]` is the log-scale-free multiplier on the odds of agreeing, so it is not
easy to read on its own. The posterior of `phi` and the ARI between the views'
partitions in each draw are more directly interpretable. The ARI is computed per
draw, from the sampled allocations, so it is unaffected by labelling.

```{r phi, eval = have_posterior}
posterior::summarise_draws(
  posterior::subset_draws(draws, variable = "phi"),
  "median", ~posterior::quantile2(.x, c(0.05, 0.95))
)
```

```{r ari-draws}
chains <- unclass(fit)[seq_along(fit)]
keep <- seq(floor(burn / 10) + 2, 301)        # saved samples after the burn in
pair_ari <- function(a, b) {
  unlist(lapply(chains, function(ch) {
    vapply(keep, function(s) ari(ch$allocations[s, , a], ch$allocations[s, , b]), numeric(1))
  }))
}
ari_post <- sapply(list("1-2" = c(1, 2), "1-3" = c(1, 3), "2-3" = c(2, 3)),
                   function(p) pair_ari(p[1], p[2]))
round(apply(ari_post, 2, quantile, c(0.05, 0.5, 0.95)), 2)
```

For comparison, the ARI between the true partitions:

```{r ari-truth}
c(true_12 = ari(z1, z2), true_13 = ari(z1, z3), true_23 = ari(z2, z3))
```

Views 1 and 2 are strongly associated, and the posterior ARI is close to the true
value; views 1 and 3 and views 2 and 3 are associated little or not at all. Note
where the prior matters: for the unrelated pairs `phi` is small but imprecise, and
it is not pushed to zero, because the prior on `phi` has little mass near zero. A
small `phi` is therefore not by itself evidence of independence, and its value there
depends on the prior. The ARI between partitions, which describes the allocation,
is the more informative summary. The sensitivity section returns to this.

### Which items fuse?

The fusion probability of an item for views $l$ and $m$ is the posterior
probability that it takes the same component in both. It is the item-level
output that a single global clustering could not give.

```{r fusion}
plot(fit, type = "fusion", burn = burn)
```

The sorted curve for views 1 and 2 has a plateau near one (items that follow the
shared structure) and a tail of items that do not. Here the truth is known, so we
can check that the tail is the set of items we broke. In real data you would look
at those items, in the measurements, for a reason.

```{r fusion-items}
pooled <- predictFromMultipleChains(fit, burn = burn)
fusion12 <- pooled$fusion_probabilities$fused_items_12
low <- fusion12 < 0.5
table(flagged_low_fusion = low, truly_broken = seq_len(N) %in% broken)
```

Some of the 18 re-drawn items kept their cluster by chance (a re-draw chooses the
same cluster a third of the time), and no method can flag these:

```{r fusion-chance}
sum(z2[broken] == z1[broken])     # re-drawn items that happen to agree
```

The flagged items are the broken ones that really differ, and the list contains no
other item.
The fusion probability for view pairs that share little structure is low across the
board, and that is information too:

```{r fusion-others}
round(sapply(pooled$fusion_probabilities, mean), 2)
```

### The partitions

A point estimate of each view's partition is the one that minimises the posterior
expected loss (variation of information by default, through `salso`). The
posterior similarity matrix, whose entries are the probability that two items share
a cluster, shows how sure that partition is. Report both: a partition without its
uncertainty overstates what the data say.

```{r partitions}
sapply(pooled$pred, function(p) length(unique(p)))
c(ARI_view1 = ari(pooled$pred[[1]], z1), ARI_view2 = ari(pooled$pred[[2]], z2),
  ARI_view3 = ari(pooled$pred[[3]], z3))
```

The number of *occupied components* in the draws is not the number of clusters.
A component holding a few items counts as occupied, so the draws can occupy more
components than the point estimate has clusters, particularly in a weakly
informative view such as view 3. Report the point estimate's clusters and the
posterior of the occupied count together.

```{r psm}
plot(fit, type = "psm", view = 1, burn = burn)
plot(fit, type = "psm", view = 3, burn = burn)
```

View 1 shows three blocks with little uncertainty. View 3 (four binary features) is
much less certain, and a weak view is exactly where the coupling helps most when it
is associated with a strong one, and where it cannot help when it is not. Here
it is not associated, so view 3 has to stand on its own data.

## Does the fitted model reproduce the data?

Posterior predictive checks compare the data with replicates drawn from the fitted
model. They find gross misfit, not small misfit: the data are used twice, so a
value near 0 or 1 is evidence of a problem and a value in the middle is not
evidence of a good model. Pick statistics that target what the model might get
wrong. Here, the spread of each continuous view and the proportion of ones in
each binary feature.

```{r ppc}
post_sims <- simulatePosteriorPredictive(fit, X, n_draws = 100, burn = burn)
sd_stat <- function(m) stats::sd(m[, 1], na.rm = TRUE)
c(view1_sd_p = predictiveCheck(X, post_sims, sd_stat, view = 1)$p_value,
  view2_sd_p = predictiveCheck(X, post_sims, sd_stat, view = 2)$p_value,
  view3_mean_p = predictiveCheck(X, post_sims, function(m) mean(m[, 1]), view = 3)$p_value)
plotPredictiveCheck(X, post_sims, view = 1, column = 1)
```

Each replicate uses the sampled allocations of the draw it came from, so these
checks test the within-cluster densities given the partition. They do not test
whether the partition itself, or the association between views, is right; the
agreement between independent chains and the sensitivity analysis below are the
checks for those.

**Prior-data conflict and the model-based comparison.** `pointwiseLogLik()` gives
the log-likelihood of each item with its assignments in every view summed out, as
the matrix that `loo` expects. That is the quantity to use to compare models that
differ in `K`, in the prior or in the component density, by approximate
leave-one-out cross-validation. For mixtures, check the Pareto $\hat k$ values: a
high value for an item means that the draws are very sensitive to it, and the
comparison for that item is not trustworthy.

```{r loo, eval = have_loo}
ll <- pointwiseLogLik(fit, burn = burn)
r_eff <- loo::relative_eff(exp(ll), chain_id = attr(ll, "chain_id"))
loo_default <- loo::loo(ll, r_eff = r_eff)
loo_default
```

## Sensitivity analysis

A result you would report should survive reasonable changes in the things you
chose. Two are worth testing here, because both matter more for MDI than for a
single mixture.

1. **The prior on `phi`.** The default has mean 10 and little mass near zero: it
   *expects* association. A weaker prior asks whether the data, not the prior,
   produce the association between views 1 and 2, and whether the apparent
   association of the unrelated pairs is the prior's doing.
2. **The upper bound `K`.** If the posterior is sensitive to `K`, the data are
   not determining the number of clusters and the prior is.

```{r sensitivity}
sens <- function(prior, K, seed) {
  set.seed(seed)
  fitMDI(X, n_chains = 2, R = 2000, thin = 10, types = types, K = K, prior = prior,
         betas = ptLadder(3, 0.25), split_merge = 2, burn = 1000, save_pointwise = TRUE,
         verbose = FALSE)
}
fit_weak <- sens(mdiPrior(mass_shape = 2, mass_rate = 0.5, phi_shape = 1, phi_rate = 0.5),
                 K, seed = 11)
fit_bigK <- sens(prior, c(10, 10, 10), seed = 12)
```

```{r sensitivity-compare, eval = have_posterior}
phi_summary <- function(f, burn) {
  d <- posterior::as_draws(f, burn = burn)
  s <- posterior::summarise_draws(posterior::subset_draws(d, variable = "phi"),
                                  "median", ~posterior::quantile2(.x, c(0.05, 0.95)))
  s$variable <- paste(s$variable)
  s
}
rbind(cbind(model = "default phi prior", phi_summary(fit, burn)),
      cbind(model = "weak phi prior", phi_summary(fit_weak, 1000)),
      cbind(model = "K = 10", phi_summary(fit_bigK, 1000)))
```

(The two sensitivity fits use two shorter chains each.) The association between views 1 and 2 stays large under all three, so that
conclusion is driven by the data. The `phi` of the unrelated pairs moves with the
prior, which is the sign that it is prior-dominated, as expected when two views
carry no shared signal: look at the partition-level summaries (the ARI between
views, the fusion probabilities), which stay the same.

```{r sensitivity-partitions}
pp_weak <- predictFromMultipleChains(fit_weak, burn = 1000)
pp_bigK <- predictFromMultipleChains(fit_bigK, burn = 1000)
data.frame(
  model = c("default phi prior, K = 6", "weak phi prior", "K = 10"),
  ARI_view1 = c(ari(pooled$pred[[1]], z1), ari(pp_weak$pred[[1]], z1), ari(pp_bigK$pred[[1]], z1)),
  ARI_view2 = c(ari(pooled$pred[[2]], z2), ari(pp_weak$pred[[2]], z2), ari(pp_bigK$pred[[2]], z2)),
  ARI_view3 = c(ari(pooled$pred[[3]], z3), ari(pp_weak$pred[[3]], z3), ari(pp_bigK$pred[[3]], z3)),
  mean_fusion_12 = c(mean(pooled$fusion_probabilities[[1]]), mean(pp_weak$fusion_probabilities[[1]]),
                     mean(pp_bigK$fusion_probabilities[[1]]))
)
```

```{r sensitivity-loo, eval = have_loo}
ll_weak <- pointwiseLogLik(fit_weak, burn = 1000)
ll_bigK <- pointwiseLogLik(fit_bigK, burn = 1000)
loo_of <- function(ll) loo::loo(ll, r_eff = loo::relative_eff(exp(ll), chain_id = attr(ll, "chain_id")))
loo::loo_compare(list(default = loo_default, weak_phi = loo_of(ll_weak), K10 = loo_of(ll_bigK)))
```

Differences in expected log predictive density of less than about two standard
errors mean the data do not clearly separate the models, and any of them can be
reported. A real
analysis would also vary the density of each view (`"G"` against `"MVN"`) and,
where outliers are expected, add the outlier component (`"TAGM"`).

## Agreement across independent runs

Convergence diagnostics test one kind of failure. The remaining one is that every
chain sits in the same wrong basin. Two checks are cheap. First, compare the
co-clustering probabilities of runs that differ in their sampler settings, here the
plain chains against the tempered chains with split-merge moves. All eight chains are
compared, with the same burn in.

```{r compare-runs}
compareRuns(list(fit_plain, fit), view = 1, burn = 1500)
```

Read the mean between-run standard deviation, and check that the pairs whose
probabilities differ by more than the tolerance are the ambiguous ones (items
between clusters), where a few hundred draws per chain leave visible Monte Carlo
variation. The plain chains include the one that sat in a different mode, which is
the kind of difference this comparison is meant to show. A systematic difference, or
one among confidently clustered items, is a warning. Second, for a result that matters,
repeat with `smcMDI()` or `smcReplicates()`, which start from the prior, not from
the chains' initial labels, and see whether they agree.

## Prediction and semi-supervised use

A fitted model gives the predictive density of new items, and their probability of
co-clustering with each fitted item, summing over the sampled weights, `phi` and
component parameters. A new item may be missing entries, or a whole view.

```{r predict}
# One new item, given as one row per view: it matches cluster 1 in view 1, is
# missing from view 2, and is partly informative in view 3
new_items <- list(
  matrix(c(-3, 0), 1),
  matrix(c(NA, NA, NA), 1),
  matrix(c(1, 1, 0, 0), 1)
)
pred <- predictMDI(fit, X, new_items, burn = burn, coclustering = TRUE)
round(pred$log_predictive, 1)
round(sort(pred$coclustering[[1]][1, ], decreasing = TRUE)[1:5], 2)
```

The second line shows the five fitted items most likely to share a cluster with the
new item in view 1 and the probabilities. They should be items of the cluster whose
mean the new item matches. If some of the items have known classes, they can be
held fixed and the model then gives class probabilities for the others. The
classes of the fixed items define the first components of that view, in sorted
order, and the labels of the other views are not constrained.

```{r semisupervised}
known <- sort(sample(N, 30))
fixed <- matrix(0, N, 3)
fixed[known, 1] <- 1
start <- matrix(1, N, 3)
start[, 1] <- z1                      # only the rows with fixed = 1 are used
ss <- fitMDI(X, n_chains = 2, R = 3000, thin = 10, types = types, K = K, prior = prior,
             initial_labels = start, fixed = fixed, burn = 1500, verbose = FALSE)
ss_pred <- predictFromMultipleChains(ss, burn = 1500)
held_out <- setdiff(seq_len(N), known)
mean(ss_pred$pred[[1]][held_out] == z1[held_out])    # accuracy on the held-out items
```

The class probabilities are in `ss_pred$allocation_probability[[1]]` (items by
components). Check them in the same way as the clusters: how well calibrated are
the high-probability predictions on items whose class you hold out?

## What to report

* The data, the density of each view and the coding of missing values.
* The priors, with the prior predictive check, and the settings: chains,
  iterations, thinning, burn in, `K` and any extra moves.
* The convergence summary of the label-invariant quantities, and which ones mix
  slowly (the concentrations) and why that does not matter for the questions.
* The partitions of each view with their uncertainty (posterior similarity), the
  posterior of the association between views (`phi` and the ARI between
  partitions), and the item-level fusion probabilities. State that fusion
  probabilities compare component indices across views.
* The posterior predictive checks and what they cannot show, and the sensitivity of
  the conclusions to the prior on `phi` and to `K`.
* What you did not check: the possibility that all chains share a wrong mode, and
  the adequacy of the component densities beyond the statistics tested.

## References

Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B., Yao, Y.,
Kennedy, L., Gabry, J., Burkner, P.-C. and Modrak, M. (2020). Bayesian workflow.
arXiv:2011.01808.

Kirk, P., Griffin, J. E., Savage, R. S., Ghahramani, Z. and Wild, D. L. (2012).
Bayesian correlated clustering to integrate multiple datasets. *Bioinformatics*,
28(24), 3290-3297.

Rousseau, J. and Mengersen, K. (2011). Asymptotic behaviour of the posterior
distribution in overfitted mixture models. *Journal of the Royal Statistical
Society B*, 73(5), 689-710.

Vehtari, A., Gelman, A., Simpson, D., Carpenter, B. and Burkner, P.-C. (2021).
Rank-normalization, folding, and localization: an improved Rhat for assessing
convergence of MCMC. *Bayesian Analysis*, 16(2), 667-718.

Vehtari, A., Gelman, A. and Gabry, J. (2017). Practical Bayesian model evaluation
using leave-one-out cross-validation and WAIC. *Statistics and Computing*, 27,
1413-1432.
